The Effect of Episodic Retrieval on Inhibition in Task Switching

### 2015-11-26 2015/Episodic Retrieval & Inhibition/Manuscript Files/main.tex
added prior robustness check

%%%--------------------------------------------------------------------------------------------
\documentclass[a4paper, man, natbib]{apa6}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes} 
\usepackage{etoolbox}

% \newtoggle{jou}
% \toggletrue{jou}
% \togglefalse{jou}

@@

\note{{\bf Word Count:} 3,124 (Main body, not including abstract or references.)}
% \note{{\bf Under Review---Please, no direct quotes.}}

@@

How the cognitive system addresses the stability--flexibility dilemma has been studied using the task switching paradigm \citep{Grange2014a,Kiesel2010,Vandierendonck2010}, wherein participants have to make speeded responses to simple cognitive tasks. For example, participants might be presented with number stimuli, and be asked to switch between judging whether the stimulus is odd/even or lower/higher than five. One cognitive process thought to be essential for task switching performance is the inhibition of competing tasks. Evidence for inhibition comes from the so-called backward inhibition paradigm \citep{Koch2010, Mayr2000}, where participants are required to switch between three tasks (A, B, and C).  It is a consistent finding that AB$A$ task switching sequences are performed slower and with less accuracy than CB$A$ sequences. This \emph{n--2 repetition cost} is thought to reflect the persisting inhibition of task A in an ABA sequence, which hinders its re-activation in the last trial of the triplet.

The n--2 repetition cost is important as it appears to be robust against non-inhibitory explanations \citep{Mayr2007}. One potential explanation of the n--2 repetition cost that does not assume inhibition---the episodic retrieval account \citep{Neill1997}---was explored by \cite{Mayr2002}. This theory suggests (adapted to a task switching context) that when a task has been performed an episodic trace of trial parameters (such as the cue, stimulus characteristics, and the response made) is stored in memory. When this task is cued again, retrieval of the most recent episodic trace occurs. If the parameters of the presented trial differ from that of the retrieved episode (e.g., a different response is required), a mismatch cost occurs relative to if the retrieved episode matches parameters of the current trial (which will prime performance). Thus, according to this account, n--2 repetition costs reflect a mismatch cost, as trial parameters typically differ between both instances of task A in an ABA sequence.

\cite{Mayr2002} assessed this alternative explanation by developing a paradigm where the match between trial parameters across an ABA sequence could be manipulated (see Figure \ref{fig:mayrExperiment}). Participants were required to apply a spatial transformation of a target's location according to one of three stimulus--response rules (vertical, horizontal, or diagonal transformations). For example, if the target is in the bottom-left of the display and the current rule is ``vertical'', the correct response would be an upper-left response (as this reflects a vertical transformation of a bottom-left target). The parameters that might be stored in an episodic trace include (at least) the cue used, the target's location, and the response made. In this paradigm, trial parameters can either match or mismatch across ABA and CBA sequences. For example, if the target is in the same location for task A across an ABA sequence, the same response will be required (i.e., an n--2 response repetition). In this case, the episodic retrieval account would predict facilitated performance relative to an n--2 response switch as in the former case the current trial parameters match perfectly the retrieved episodic trace. Thus, the episodic account would predict reduced (or zero) n--2 task repetition costs for n--2 response repetitions compared to n--2 response switches (where there is a mismatch between trial parameters and the retrieved episodic trace).

@@

The Bayes factor for the critical statistical test from Mayr's (2002) study ($BF_{10}$ = 0.315) suggests that the null hypothesis is about 3 times more likely than the alternative, given the data, which provides anecdotal--moderate support for the null. As this Bayesian analysis does not provide decisive evidence for the null, and given the importance of assessing to what extent (if any) episodic retrieval can influence n--2 task repetition costs in task switching, we were interested in performing a replication of Mayr's study to gather more evidence.

The current study replicated critical aspects of Mayr's (2002) design relevant to n--2 task repetitions. As Mayr's study was also interested in exploring to what extent the paradigm generated typical task switching effects (such as switch costs, effects of preparation, and effects of the delay between tasks), we made some changes to the original design of Mayr. Specifically, immediate task repetitions were not allowed; it has been shown that including immediate task repetitions reduces the n--2 task repetition cost \citep{Philipp2006}, so we removed this possibility. We also did not manipulate preparation via the cue--stimulus interval, as Mayr reported no effect of this variable on n--2 task repetition costs. Mayr did find that n--2 repetition costs were larger with shorter response--cue intervals (RCI; the time between the response to one trial and the onset of the cue for the next trial), a typical finding for n--2 task repetition costs \citep{Gade2005, Grange2009, Mayr2000}. Therefore, in order to obtain large n--2 task repetition costs we only used a short RCI.

@@

\subsection{Procedure}
On each trial a cue was presented above the frame for 150ms. After this time, the circle stimulus appeared in one of the four corners of the frame. The cue remained on the screen during stimulus presentation. The task required participants to mentally make a spatial transformation of the stimulus' location according to the rule dictated by the task cue, and to make a spatially-congruent response to this translated location. For example, if the stimulus was in the top-right corner of the frame, the participant would need to make a top-left response if the task was ``horizontal'', a bottom-right response if the task was ``vertical'', and a bottom-left response if the task was ``diagonal''. Responses were made on the numerical part of the keyboard, using the ``4'', ``5'', ``1'', and ``2'' keys for top-left, top-right, bottom-left, and bottom-right responses respectively. Participants were asked to respond as quickly and as accurately as possible. Once a response was registered, the frame was cleared for a response--cue interval of 150ms, after which time a new cue appeared for the next trial. The cue for the next trial was chosen randomly with the constraint that task repetitions could not occur (cf., Mayr, 2002); the stimulus location was chosen randomly with no constraints on each trial. If participants made an error, the word ``Error!'' appeared in red for 1000ms above the stimulus frame (in the cue's position).

@@

@@

Our results differ to those reported by \cite{Mayr2002}. We suggest that our modifications to Mayr's original design improved the sensitivity of the paradigm to be able to detect potential episodic retrieval effects on n--2 task repetitions. The modifications were made to enhance the magnitude of n--2 task repetition costs, so that modulation of this cost with response repetitions could be more easily detected. Specifically, removing the possibility of immediate task repetitions likely enhanced the n--2 task repetition cost as it has been shown that allowing immediate repetitions reduces this cost \citep{Philipp2006}. We also used the shortest RCI used by Mayr (150ms), as it is a consistent finding that this increases the n--2 repetition cost \cite{Gade2005, Grange2009, Mayr2000, Mayr2002}. Any one of these modifications could explain the difference in results between this study and that of Mayr (2002). Future research should aim to examine the potential role for episodic retrieval influencing n--2 task repetition costs by providing further replications to this study \citep{OpenScienceCollaboration2015}.

>>>>

%%%--------------------------------------------------------------------------------------------
\documentclass[a4paper, doc, natbib]{apa6}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes} 
\usepackage{etoolbox}

@@

\note{{\bf Word Count:} 3,280 (Main body, not including abstract or references.)}
% \note{{\bf Under Review---Please, no direct quotes.}}

@@

How the cognitive system solves the stability--flexibility dilemma has been studied using the task switching paradigm \citep{Grange2014a,Kiesel2010,Vandierendonck2010}, wherein participants have to make speeded responses to simple cognitive tasks. For example, participants might be presented with number stimuli, and be asked to switch between judging whether the stimulus is odd/even or lower/higher than five. One cognitive process thought to be essential for task switching performance is the inhibition of competing tasks. Evidence for inhibition comes from the so-called backward inhibition paradigm \citep{Koch2010, Mayr2000}, where participants are required to switch between three tasks (A, B, and C).  It is a consistent finding that AB$A$ task switching sequences are performed slower and with less accuracy than CB$A$ sequences. This \emph{n--2 task repetition cost} is thought to reflect the persisting inhibition of task A in an ABA sequence, which hinders its re-activation in the last trial of the triplet.

The n--2 task repetition cost is important as it appears to be robust against non-inhibitory explanations \citep{Mayr2007}. One potential explanation of the n--2 task repetition cost that does not assume inhibition---the episodic retrieval account \citep{Neill1997}---was explored by \cite{Mayr2002}. This theory suggests (adapted to a task switching context) that when a task has been performed an episodic trace of trial parameters (such as the cue, stimulus characteristics, and the response made) is stored in memory. When this task is cued again, retrieval of the most recent episodic trace occurs. If the parameters of the presented trial differ from that of the retrieved episode (e.g., a different response is required), a mismatch cost occurs relative to if the retrieved episode matches parameters of the current trial (which will prime performance). Thus, according to this account, n--2 task repetition costs reflect a mismatch cost, as trial parameters typically differ between both instances of task A in an ABA sequence.

\cite{Mayr2002} assessed this alternative explanation by developing a paradigm where the match between trial parameters across an ABA sequence could be manipulated (see Figure \ref{fig:mayrExperiment}). Participants were required to apply a spatial transformation of a target's location according to one of three stimulus--response rules (vertical, horizontal, or diagonal transformations). For example, if the target is in the bottom-left of the display and the current rule is ``vertical'', the correct response would be an upper-left response (as this reflects a vertical transformation of a bottom-left target). The parameters that might be stored in an episodic trace include (at least) the cue presented, the target's location, and the response made. In this paradigm, trial parameters can either match or mismatch across ABA and CBA sequences\footnote{Note that with one cue per task, the cue for task A will always be the same across an ABA sequence.}. For example, if the target is in the same location for task A across an ABA sequence, the same response will be required (i.e., an n--2 response repetition). In this case, the episodic retrieval account would predict facilitated performance relative to an n--2 response switch as in the former case the current trial parameters match perfectly the retrieved episodic trace. Thus, the episodic account would predict reduced (or zero) n--2 task repetition costs for n--2 response repetitions compared to n--2 response switches (where there is a mismatch between trial parameters and the retrieved episodic trace).

@@

The Bayes factor for the critical statistical test from Mayr's (2002) study ($BF_{10}$ = 0.315)\footnote{For all Bayes factors reported in this paper we use a non-informative prior on the alternative hypothesis, which is assumed to be distributed as a Cauchy with scaling factor $r$ = 0.707.} suggests that the null hypothesis is about 3 times more likely than the alternative, given the data, which provides anecdotal-to-moderate support for the null. As this Bayesian analysis does not provide decisive evidence for the null, and given the importance of assessing to what extent (if any) episodic retrieval can influence n--2 task repetition costs in task switching, we were interested in performing a replication of Mayr's study to gather more evidence.

The current study replicated critical aspects of Mayr's (2002) design relevant to n--2 task repetitions. As Mayr's study was also interested in exploring to what extent the rule-switching paradigm generated typical task switching effects (such as switch costs, effects of preparation, and effects of the delay between tasks), we made some changes to the original design of Mayr. Specifically, immediate task repetitions were not allowed; it has been shown that including immediate task repetitions reduces the n--2 task repetition cost \citep{Philipp2006}, so we removed this possibility. We also did not manipulate preparation via the cue--stimulus interval, as Mayr reported no effect of this variable on n--2 task repetition costs. Mayr did find that n--2 repetition costs were larger with shorter response--cue intervals (RCI; the time between the response to one trial and the onset of the cue for the next trial), a typical finding for n--2 task repetition costs \citep{Gade2005, Grange2009, Mayr2000}. Therefore, in order to obtain large n--2 task repetition costs we only used a short RCI.

@@

\subsection{Procedure}
On each trial a cue was presented above the frame for 150ms. After this time, the circle stimulus appeared in one of the four corners of the frame. The cue remained on the screen during stimulus presentation. The task required participants to mentally make a spatial transformation of the stimulus' location according to the rule dictated by the task cue, and to make a spatially-congruent response to this translated location. For example, if the stimulus was in the top-right corner of the frame, the participant would need to make a top-left response if the task was ``horizontal'', a bottom-right response if the task was ``vertical'', and a bottom-left response if the task was ``diagonal''. Responses were made on the numerical part of the keyboard, using the ``4'', ``5'', ``1'', and ``2'' keys for top-left, top-right, bottom-left, and bottom-right responses respectively. Participants were asked to respond as quickly and as accurately as possible, using the index finger of their right hand. In between trials, participants were asked to place their finger in the centre of the four response keys. Once a response was registered, the frame was cleared for a response--cue interval of 150ms, after which time a new cue appeared for the next trial. The cue for the next trial was chosen randomly with the constraint that task repetitions could not occur (cf., Mayr, 2002); the stimulus location was chosen randomly with no constraints on each trial. If participants made an error, the word ``Error!'' appeared in red for 1000ms above the stimulus frame (in the cue's position).

@@

\subsubsection{Robustness Check on Bayesian Priors}
We used a default prior distribution on the effect size for the alternative hypothesis in the Bayes factor analysis, which was assumed to be distributed as a Cauchy distribution with scale parameter $r$ = 0.707. To examine how robust our findings were to the choice of prior, we conducted a robustness check by re-calculating the Bayes factor across a wide range of priors by varying scale parameter $r$: As $r$ increases, the prior distribution becomes wider, reflecting a prior for larger effects; $r$s closer to zero make the distribution narrower, reflecting a prior for smaller effects. The robustness check is plotted in Figure \ref{fig:robustPrior}. This figure shows our conclusions (and our stopping rule) are robust to the prior used: Evidence favours the alternative hypothesis across all prior widths.

\begin{figure*}
\begin{center}
\includegraphics[width = \textwidth]{Images/robustPrior.pdf}
\caption{Robustness check of Bayes factor across a wide range of priors.}
\label{fig:robustPrior}
\end{center}
\end{figure*}

@@

Our results differ to those reported by \cite{Mayr2002}. We suggest that our modifications to Mayr's original design improved the sensitivity of the paradigm to be able to detect potential episodic retrieval effects on n--2 task repetitions. The modifications were made to enhance the magnitude of n--2 task repetition costs, so that modulation of this cost with response repetitions could be more easily detected. Specifically, removing the possibility of immediate task repetitions likely enhanced the n--2 task repetition cost as it has been shown that allowing immediate repetitions reduces this cost \citep{Philipp2006}. We also used the shortest RCI used by Mayr (150ms), as it is a consistent finding that this increases the n--2 repetition cost \cite{Gade2005, Grange2009, Mayr2000, Mayr2002}. Any one of these modifications could explain the difference in results between this study and that of Mayr (2002). Future research should aim to examine the potential role for episodic retrieval influencing n--2 task repetition costs by providing further replications and conceptual extensions to this study \citep{OpenScienceCollaboration2015}.
